Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-05 WSSSPE2_report/WSSSPE2_review.tex
Clarify LBNL.
Section 8: various minor changes.

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are
(from a Git or other repository), then allows the user to manually edit the output to
add/remove people, define roles.
\item Work with repositories to encourage them to provide the information we
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption with a substantial scientific organization such as LBNL (Lawrence Berkeley National Laboratory).
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software
we use, and encourage the software developers to make their software citable
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat
standardized so that they can be imported into papers.  Also make sure that
standard LaTeX style files understand and accept these entries.
\end{enumerate}

@@

The group defined the following actions to be performed:
\begin{enumerate}
\item Investigate existing taxonomies for roles and contributorships.
\item Investigate prototype badging initiatives.
\item Investigate journals focusing on publishing peer-reviewed research
software.
\item Investigate guidelines and checklists of best practices.
\item Communicate the results of the above investigations to the WSSSPE
community, and to decision-making bodies (funders, publishers, universities,
and tenure committee representatives).
\item Ensure engagement of the broader research community in this discussion.
\end{enumerate}

@@

Prizes in different categories were discussed (like Academy Awards), for
example: best contribution (non-founder), broadest diversity of contributions,
best tutorials or documentation, best leadership transition (award ex leader and
new leader), best generalization (taking something that was limited and making
it more general), and best mentorship of contributors (bringing others into the
community).

Other (non-prize) forms of incentives are
\begin{itemize}
\item Converting reputation by joining the ASF, Google, etc.
\item Inviting ASF and open source people to contribute to scientific code (at
least on questions of overlapping
\item Template for assessing scientific contributions made through software.
\end{itemize}

The group suggested that to determine who should give out prizes, perhaps we
should find those who we think would be awarded prizes, then ask them who they
would want to receive a prize from, then ask those organizations to see if they
are willing to be involved in the process.

One of the group's ideas was to create a funding program for disciplines or
other organizations to create a prize program. We would provide a framework and
a set of requirements, for example: awards to individuals; must award in 5 or 6
areas; must have a jury and/or objective criteria that includes senior/junior
domain people and technology people; must have the recipients awarded at a
relevant event; and must provide citations that explain why the prizes should be
awarded. Different organizations could then decide to sign up to the framework
and give out awards under this general brand. However, there was a concern that
having many organizations award prizes may reduce the impact of the prizes.

Possible groups that might give out awards, either under our framework, or more
generally, are AAAS, Nature Publishing Group, ACM/IEEE, Astro, Ecology Society
of America, etc. Perhaps this could be a joint technology/science partnership,
for example, the [Apache $|$ Mozilla]-[AAAS $|$ disciplinary society] prize.

@@

The group discussion if there should be different criteria for ``established''
members of the community versus junior members, and if prizes should be
restricted to junior members, but left these as open questions.

>>>>

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are
(from a Git or other repository), then allows the user to manually edit the output to
add/remove people, define roles.
\item Work with repositories to encourage them to provide the information we
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption with a substantial scientific organization such as the Lawrence Berkeley National Laboratory (LBNL).
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software
we use, and encourage the software developers to make their software citable
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat
standardized so that they can be imported into papers.  Also make sure that
standard LaTeX style files understand and accept these entries.
\end{enumerate}

@@

The group defined the following actions to be performed:
\begin{enumerate}
\item Investigate existing taxonomies for roles and contributorships.
\item Investigate prototype badging initiatives.
\item Investigate journals focusing on publishing peer-reviewed research
software.
\item Investigate guidelines and checklists of best practices.
\item Communicate the results of the above investigations to the WSSSPE
community and decision-making bodies (funders, publishers, universities,
and tenure committee representatives).
\item Ensure engagement of the broader research community in this discussion.
\end{enumerate}

@@

Prizes in different categories were discussed (like Academy Awards), for
example: best contribution (non-founder), broadest diversity of contributions,
best tutorials or documentation, best leadership transition (award ex-leader and
new leader), best generalization (taking something that was limited and making
it more general), and best mentorship of contributors (bringing others into the
community).

Other (non-prize) forms of incentives are
\begin{itemize}
\item Converting reputation by joining the ASF, Google, etc.
\item Inviting ASF and open source people to contribute to scientific code (at
least on overlapping questions)
\item Template for assessing scientific contributions made through software.
\end{itemize}

The group suggested that to determine who should give out prizes, perhaps we
should find those who we think would be awarded prizes, then ask them who they
would want to receive a prize from, and subsequently contact those organizations to see if they
are willing to be involved in the process.

One of the groups's ideas was to create a funding program for disciplines or
other organizations to create a prize program. We would provide a framework and
a set of requirements, for example: awards to individuals; must award in 5 or 6
areas; must have objective criteria and a jury that includes senior/junior
domain experts and technologists; must have the recipients awarded at a
relevant event; and must provide citations that explain why the prizes should be
awarded. Different organizations could then decide to sign up to the framework
and give out awards under this general brand. However, there was a concern that
having many organizations award prizes may reduce the impact of the prizes.

Possible groups that might give out awards, either under our framework, or more
generally, are AAAS, Nature Publishing Group, ACM, IEEE, Astro, Ecology Society
of America, etc. Perhaps this could be a joint technology/science partnership,
for example, the [Apache $|$ Mozilla]-[AAAS $|$ disciplinary society] prize.

@@

The group discussed if there should be different criteria for ``established''
members of the community versus junior members, and if prizes should be
restricted to junior members, but left these as open questions.
